Plotting the Single-Electron Solution to Schr\"{o}dinger's Equation with OpenCL

### 2010-05-06 paper/paper.tex
More data, more plots, etc.

@@

@@

To keep the implementation of this algorithm within the scope of this project, we also decided to use a fixed-camera orthogonal projection. This worked to significantly simplify the process of iteration over all of the samples, as with this restriction, the program simply has to iterate over all of the pixels in the image, evaluating $n$ points in the $z$ dimension for each pixel, without worrying about complex transformations between coordinate systems.

@@

\subsection{GPUs}

@@

\item NVIDIA GeForce 8600M GT, 32$\times$475MHz, 512MB, ???\$\footref{fn:prices}\footnote{Tested on Mac OS X, with Apple OpenCL\label{fn:osx}}

@@

\item Intel Core i7 620M, 2$\times$3333MHz, 4GB, 332\$\footref{fn:prices}\footref{fn:osx}

@@

\begin{figure}
    \includegraphics[width=84.5mm]{speedupPlot.pdf}
    \caption{Speedup of fastest GPU over fastest CPU as resolution increases; larger values are better}
    \label{fig:speedupPlot}
\end{figure}

As one can see in \ref{fig:runtimePlot}, both GPUs significantly outperform all of the CPUs

@@

>>>>

All of the equations below are borrowed from \cite{quantumBook}, but have been slightly modified to fit our purposes, as discussed in section \ref{simpSection}.

@@

\label{simpSection}

@@

To keep the implementation of this algorithm within the scope of this project, we also decided to use a fixed-camera orthogonal projection. This worked to significantly simplify the process of iteration over all of the samples, as with this restriction, the program simply has to iterate over all of the pixels in the image, evaluating many points in the $z$ dimension for each pixel, without worrying about complex transformations between coordinate systems.

\subsection{Evaluation}

For each pixel in the image, we spawn an instance of our OpenCL kernel which evaluates 2000 samples in the $z$ dimension. Each sample represents the value of the density function, $\psi$, at that point in space (the cartesian coordinates of the image have to be converted into spherical coordinates in order to evaluate $\psi$). All 2000 samples are summed and stored into their respective pixel in the image buffer.

Once computation has completed, the buffer is copied back from video memory to system memory (or, if the program was running on the CPU, this step is skipped --- this is the primary overhead incurred by the use of GPU computation). There, pixel values are scaled linearly so that the brightest pixel is 100\% white.

@@

\subsection{GPU}

@@

@@

\item Intel Core i7 620M, 2$\times$3333MHz, 4GB, 332\$\footref{fn:prices}\footnote{Tested on Mac OS X, with Apple OpenCL\label{fn:osx}}

@@

\begin{figure}
    \includegraphics[width=84.5mm]{speedupPlot.pdf}
    \caption{Speedup of fastest GPU (4890) over fastest CPU (Core 2 Quad) as resolution increases; larger values are better}
    \label{fig:speedupPlot}
\end{figure}

As one can see in figure \ref{fig:runtimePlot}, the GPU significantly outperforms all of the CPUs. This is not surprising, as it can perform more than an order of magnitude more parallel computations, but it does validate the premise of this experiment.

@@

\begin{figure}
    \includegraphics[width=84.5mm]{dollarPlot.pdf}
    \caption{Samples per second per dollar with a fixed resolution ($500\times500$); larger values are better}
    \label{fig:costEffectiveness}
\end{figure}

\begin{figure}
    \includegraphics[width=84.5mm]{transistorPlot.pdf}
    \caption{Samples per second per million transistors with a fixed resolution ($500\times500$); larger values are better}
    \label{fig:transistorPlot}
\end{figure}

\begin{figure}
    \includegraphics[width=84.5mm]{wattPlot.pdf}
    \caption{Samples per second per heavy-load watt with a fixed resolution ($500\times500$); larger values are better}
    \label{fig:wattPlot}
\end{figure}
